---
title: "CUNY DATA 607_Wk2_Herold -- Integrating R and SQL"
output:
  html_document:
    df_print: paged
---

#### Assignment for Week 2: 

#### Choose six recent popular movies. Ask at least five people that you know (friends, family, classmates, imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5. Take the results (observations) and store them in a SQL database. Load the information into an R dataframe.


##### The following packages are loaded first:

##### - RMySQL
##### - sqldf
##### - htmlTable
##### - downloader
##### - reshape2
##### - ggplot2



```{r packages, ECHO = TRUE, include=FALSE, cache=FALSE}
require(RMySQL)
require(dplyr)
require(sqldf)
require(ggplot2)
require(downloader)
require(htmlTable)
require(reshape2)
require(ggplot2)
```



To load the SQL data into R, the dbConnect and dbGetQuery functions are used from the RMySQL package. 

A variable, con, set as the DBIConnection object, is returned by dbConnect(). The SQL query language is introduced within the dbGetQuery function. 

```{r loading SQL data, ECHO = TRUE }
dbname <- "videos"
dbuser <- "root"
dbpass <- "elmstreet101"
dbhost <- "localhost"
dbport <- 3306
con <- dbConnect(RMySQL::MySQL(), host=dbhost, port=dbport, dbname=dbname, user=dbuser, password=dbpass)

movie_rate <- dbGetQuery(con, "select critic_ID, critic_name, 1_WW as WonderWoman, 2_BR as BladeRunner2049, 3_PA as War4PlanetApes, 4_AM as AmericanMade, 5_LJ as LastJedi, 6_IT as IT from critic_rate")
head(movie_rate)
```



Numeric summary is provided here:	

```{r summary}
movie_rate2 <- movie_rate
summary(movie_rate2)
```



A horizontal bar graph shows the frequency of the ratings for each movie as follows:

```{r bar graph}
bwidth <- 0.5
df <- data.frame(WW = movie_rate2$WonderWoman, BR = movie_rate2$BladeRunner2049, PA = movie_rate2$War4PlanetApes, AM = movie_rate2$AmericanMade, LJ = movie_rate2$LastJedi, IT = movie_rate$IT)
ggplot(melt(df), aes(value, fill = variable)) + coord_flip() + geom_histogram(binwidth=bwidth, position = "dodge") + xlab("Movie Ratings") + ylab("Freq") + ggtitle("Movies of 2017")
```

```{r}
dbDisconnect(con)
```


Disconnecting from MySQL.




