---
title: "CUNY DATA607_Wk3_Herold_Regex"
output: html_document
---

## Automated Data Collection with R  (p. 217) 8.3

Using Stringr package. 


```{r}
library(stringr)
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev.Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson,Homer5553642Dr. Julius Hibbert"

name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))

# Add spaces after periods and commas to later help with extracting
name <- str_replace(name, pattern = "\\.", replacement = ". ")
name <- str_replace(name, pattern = ",", replacement = ", ")

name
```
```{r}
df <- data.frame(name = name, stringsAsFactors = F)
df
```


Determining if the name has a title.

```{r}
## Removing the presumed titles from the names
df$temp.name <- str_remove(name, "[[:alpha:]]{2,}\\.")

## Check for periods after 2+ letters to signal titles
has.title <- str_detect(name, "[[:alpha:]]{2,}\\.")

## Add has.title column to dataframe
df <- data.frame(df, has.title = has.title)

df
```

To separate the first and last names, we need to detect if there are any commas, which would change the regular order of First, then Last, name.



```{r separation of names with commas}
df$has.comma <- NULL

## Check for commas for last names first
df$has.comma <- str_detect(df$temp.name, ",")

df
```

Now, we extract out the parts of temp.name, filling into the first_name and last_name fields depending on the Boolean of whether or not there was a comma in the name. Last name first. 

```{r}
df$last_name <- NULL
df$last_name[df$has.comma == TRUE] <- unlist(str_extract_all(df$temp.name[df$has.comma == TRUE], "^[[:alpha:]]{2,}"))
df$last_name[df$has.comma == FALSE] <- unlist(str_extract_all(df$temp.name[df$has.comma == FALSE], "[[:alpha:]]{2,}$"))
df$last_name
```

Then first names, after they have been padded. 

```{r}
## Need to remove padding of temp.names first
df$temp.name <- str_trim(df$temp.name, side = "both")

df$first_name <- NULL
df$first_name[df$has.comma == TRUE] <- unlist(str_extract_all(df$temp.name[df$has.comma == TRUE], "[[:alpha:][.][:blank:]]{2,}$"))
df$first_name[df$has.comma == FALSE] <- unlist(str_extract_all(df$temp.name[df$has.comma == FALSE], "^[[:alpha:]]{2,}"))

##  I recognize that I did not generalize in the period issue in the first name, coding for C. Montgomery in this problem. I think of the problems that "St." and "Jr." must cause. 
df2 <- data.frame(df$name,df$first_name,df$last_name,df$has.title)
df2

```

Trimming the first names again, then detecting for spaces to indicate two names. 


```{r}
df$first_name <- unlist(str_trim(df$first_name, side = "both"))
df2$is.twonames <- unlist(str_detect(df$first_name, " "))
df2

```

Another problem would be determining if, given three names, a second name should be part of the first or last.


## Automated Data Collection with R  (p. 217) 8.4

## [0-9]+\\$

```{r}
rawdata1 <- c("999$", "2222$", "333")
unlist(str_extract_all(rawdata1, "[0-9]+\\$"))
```

## \\b[a-z]{1,4}\\b

```{r}
rawdata2 <- c("man","bird","Way")
unlist(str_extract_all(rawdata2, "\\b[a-z]{1,4}\\b"))
```

## .*?\\.txt$

```{r}
rawdata3 <- c(".txt","wow.dog.txt", "tree.look.txt2")
unlist(str_extract_all(rawdata3, ".*?\\.txt$"))
```

## \\d{2}/\\d{2}/\\d{4}

```{r}
rawdata4 <- c("22/09/1976","65/33/9999", "653.33/8888")
unlist(str_extract_all(rawdata4, "\\d{2}/\\d{2}/\\d{4}"))
```

## <(.+?)>.+?</\\1>

```{r}
rawdata5 <- c("<d> </d>","<meta> weep </meta>","giant")
unlist(str_extract_all(rawdata5, "<(.+?)>.+?</\\1>"))
```






