---
title: "CUNY DATA607_Wk3_Herold_Chess"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project 1: Wrangling data from a Chess Tournament Cross Table
## Zachary Herold, Submitted 9/20/18


```{r}
library(stringr)

chess <- NULL
opponents.rating.matrix <- NULL
```

Loading the Chess Tournament Cross Table from the Raw Github URL. Removing the dash lines, with solution by redmode. See: https://stackoverflow.com/questions/21114598/importing-a-text-file-into-r


```{r}
url <- "https://raw.githubusercontent.com/ZacharyHerold/chinafundnews/master/tournamentinfo.txt"

lines <- readLines(url)

lines <- sapply(lines, gsub, pattern="[-]{2,}|[|]", replacement="")
lines <- c(lines[2], lines[lines!="" & lines!=lines[2]])

head(lines)

```

Extracting out the names with alphabetic characters, allowing for multiple names. 

```{r}
name <- unlist(str_extract_all(lines, "[[:alpha:]]{2,}([[:blank:]][[:alpha:]]{1,}){1,}"))

name <- name[-c(1,2)]
name
```

Trimming the lines and taking the first two characters, then subsetting out the strings of two alphabetic characters, indicating the state abbreviations. As with name, removing the first two rows.

```{r}
trim_lines <- str_trim(lines, side = "both")
state_collect <-  str_sub(trim_lines, start = 1, end = 2)
state <- unlist(str_extract_all(state_collect, "[[:alpha:]]{2}"))
state <- state[-c(1,2)]
state


```

Checking the length of the state vector.

```{r}
length(state)
```

Separating out the digits divided by a decimal point. 

```{r}
points <- unlist(str_extract_all(lines, "[[:digit:]]\\.[[:digit:]]"))
```

```{r}
length(points)
```


The "R: " characters indicate the player's ratings. 

```{r}
pre_ratings <- unlist(str_extract_all(lines, "R:( ){1,2}[[:digit:]]{3,4}"))
pre_ratings <- unlist(str_extract_all(pre_ratings, "[[:digit:]]{3,4}"))
pre_ratings
```

```{r}
length(pre_ratings)
```

COnstructing the chess dataframe.

```{r}
chess <- data.frame(name, ST = state, points, pre_ratings, stringsAsFactors = FALSE)
head(chess)
```

Retrieving the contest outcome and opponent., looking for Wins ("W"), Losses ("L"), or Draws ("D").

```{r}
results <- unlist(str_extract_all(lines, "[WLD][[:blank:]]{2,3}[[:digit:]]{0,2}"))
results
```

With 64 players and 7 rounds, there should be 448 total scores. Checking the data, should remove the "W" followed by blanks with no opponent number.

```{r}
length(results)

64 * 7
```

```{r}
results.raw <- subset(results, results != "W   ")
length(results.raw)
```

Just one short. Finding the abberation. The one "B" in line 75.

```{r}
lines[75:77]
```


```{r}
results <- unlist(str_extract_all(lines, "[WLDUH][[:blank:]]{2,3}[[:digit:]]{0,2}"))
results.raw <- subset(results, results != "W   ")
results.raw
```


```{r}
length(results.raw)
```

Manually discovering that the extra "B" is in position 408, then inserting it. Would be nice to automate this process somehow. 

```{r}
results.raw <- c(results.raw[1:407],"B", results.raw[408:length(results.raw)])
length(results.raw) 

```

From the results vector, removing the result so that only the opponent number remains. 

```{r}
opponents <- NULL

opponents <- ifelse(str_detect(results.raw, "[[:digit:]]{1,2}"), unlist(str_extract_all(results.raw, "[[:digit:]]{1,2}")), NA)
opponents

```


Checking the class of the variable opponents. Finding it is a character, converting that to a numeric.


```{r}
class(opponents)
opponents <- as.numeric(opponents)
class(opponents)
```

Creating an opponents matrix, with 64 rows for the players and 7 columns for the rounds.

```{r}
opponents.mat <- matrix(opponents, nrow = 64, ncol = 7, byrow = TRUE)
opponents.mat
```

Replacing the player ID with their ratings, again ensuring it is a numeric value. The matrix is dissolved. 

```{r}
opponents.rating <- as.numeric(pre_ratings[opponents])
opponents.rating
```

Restructuring the opponent ratings into a matrix. 

```{r}
opponents.rating.matrix <- matrix(opponents.rating, nrow = 64, ncol = 7, byrow = TRUE)
opponents.rating.matrix
```

Taking the mean of the matrix to find the average rating of opponents, after removing the NA values.

```{r}
chess$opponents.rating <- round(rowMeans(opponents.rating.matrix, na.rm = T, dims = 1))
chess
```

Finally, capturing the data into a csv file. 

```{r}
write.csv(chess, file = "chess.csv")
```


